Do Drivers Dream of Walking? An Investigation of Travel Mode Dissonance from the Perspective of Affective Values

### 2021-01-14 Dissonance_Santiago_v3/Dissonance_Santiago_v3.Rmd
v3 Revisions after comments from anonymous reviewers, Journal of Transport and Health

---
title: Do Drivers Dream of Walking? An Investigation of Travel Mode Dissonance from the Perspective of Affective Values
author:
  - name: Author 1
    email: a1@university.com
    affiliation: Some University
    footnote: Corresponding Author
  - name: Author 2
    email: a2@school.com
    affiliation: Some School
address:
  - code: Some University
    address: Department, Street, City, State, Zip
  - code: Some School
    address: Department, Street, City, State, Zip
abstract: |
   *Introduction*  
   Subjective wellbeing is a topic that has attracted considerable attention in the transportation literature in recent years. As a result, there is a burgeoning literature that investigates the impacts of travel on subjective wellbeing, and how wellbeing, in turn, can influence behavior. An important aspect of subjective wellbeing are the affective reactions of people to their experiences.   
   *Objective*  
   The objective of this paper is to analyze the affective reactions of travelers with respect to various modes of transportation. In particular, we are interested in the potential for dissonance between primary mode of travel and the mode(s) of travel identified as evoking various affective reactions.  
   *Materials and Methods*  
   The study is based on data collected from a sample of travelers in the city of Santiago, in Chile. Participants in the study were asked about their usual mode of travel, and then were asked to name their ideal mode(s) of transportation from the perspective of various affective reactions. The reactions we investigate are associated with the values of freedom, enjoyment, happiness, poverty, luxury, and status. Analysis is based on tests of independence and visualization techniques.  
   *Results*  
   The results indicate that users of public transportation experience the most dissonance in terms of affective reactions, and active travelers the least. For those travelers who experience dissonance, active travel is the mode most commonly associated with freedom, enjoyment, and happiness, whereas public transportation is most commonly associated with poverty. The automobile, in contrast, is the mode most commonly associated with luxury and status.

@@

# The master dataframe is processed for analysis. We plan to conduct two distinct types of analysis:
#
# 1. Who experiences dissonace?
#
# Analysis of dissonace using the characteristics of respondents (i.e., AGE, EDUCATION, INCOME, (Mode) USED, and TRAVEL_TIME) by ATTRIBUTE (i.e., FREEDOM et al.). In this case, dissonance is defined as follows:
#
# >> If the mode used is identical to the mode that the respondent associates with ATTRIBUTE (i.e. FREEDOM et al.), we consider that there is no dissonance; otherwise, we consider that there is dissonance. For example, if the respondent's main mode of transportation is Car, and the respondent associates Car with feelings of FREEDOM, then there is no dissonance. On the other hand, if the respondent's main mode of transportation is Car, but associates Walking with feelings of FREEDOM, there is dissonance.
# 
# Given this, we will calculate how frequently there is dissonance at different levels of AGE, EDUCATION, INCOME, MODE, TRAVEL_TIME. Tests of independence can be implemented.
#
# 2. For those respondents who experience dissonance, what mode(s) are better aligned with various affective values
#
# Analysis of dissonance using the mode USED and the mode respondents associated with ATTRIBUTE, then stratified by characteristics of respondents (i.e., AGE et al.) in this case, we see how frequently each mode was associated with ATTRIBUTE, given the main mode of transportation of the respondents.
#
# Here the dataframes are prepared to support these two types of analysis

@@

  ###################################################################################
#                                                                                 #
# Function created by Dr Paul Williamson, Dept. of Geography and Planning,        #
# School of Environmental Sciences, University of Liverpool, UK.                  #
#                                                                                 #
# Adapted from the function ctab() in the catspec packge.                         #
#                                                                                 #
# Version: 12th July 2013                                                         #
#                                                                                 #
# Output best viewed using the companion function print.crosstab()                #
#                                                                                 #
###################################################################################

@@

#If no type supplied, default to 'frequency + total' for univariate tables and to
#'frequency' for multi-dimenstional tables

@@

#If dec.places not set by user, set to 2 unlesss only one table of type frequency requested,
#in which case set to 0.  [Leaves user with possibility of having frequency tables with > 0 dp]
if (is.null(dec.places)) {
  if ((length(type)==1) & (type[1]=="frequency")) {
    dec.places <- 0
  } else {
    dec.places <-2
  }
}

@@

#If subtotals not required, and total vars > 2, create dataframe version of table, with relevent
#subtotal rows / cols dropped [Subtotals only present in tables with > 2 cross-classified vars]
t1 <- NULL
if ( (subtotals==FALSE) & (n.vars>2) )  {

@@

  #Remove backslashes from category names AND colnames
  t1 <- apply(t1[,],2, function(x) gsub("\"","",x))
  #Remove preceding and trailing spaces from category names to enable accurate capture of 'sum' rows/cols
  #[Use of grep might extrac category labels with 'sum' as part of a longer one or two word string...]
  t1 <- apply(t1,2,function(x) gsub("[[:space:]]*$","",gsub("^[[:space:]]*","",x)))

@@

A definition of subjective wellbeing is as “[g]ood mental states, including all of the various evaluations, positive and negative, that people make of their lives, and the affective reactions of people to their experiences” [@oecd2013oecd, p. 10]. As memorably put by Steg [-@Steg2005car]: is the use of a car a must, or is it a lust? A close alignment, or consonance, between affective reactions and the mode of transportation used to travel can result in greater subjective wellbeing and increase the probability of choosing that mode (i.e., by satisfying the _lust_); on the other hand, dissonance, namely the lack of concordance between the mode used and the affective reaction to that mode, would be detrimental to subjective wellbeing, especially to those for whom the use of the mode is a _must_ [@Devos2019satisfaction; @Mokhtarian2018travel]. With these considerations in mind, the present research aims to investigate _who_ experiences affective dissonance with respect to their primary mode of travel. We focus our investigation on reactions from the perspective of feelings of freedom, enjoyment, happiness, poverty, luxury, and status - concepts termed _affective values_ hereafter. Furthermore, it is possible, even if use of a particular mode is a _must_, that travelers may still _lust_ for something else - in other words, the grass may look greener from the window of a car. For this reason, we also aim to investigate which mode or modes are most commonly identified as ideal from the perspective of the aforementioned affective values.

Our research is based on data collected from a sample of travelers in the city of Santiago in Chile. The analysis sheds light on the affective reactions of people towards their primary mode of travel, and their perceptions towards 'ideal modes'. This reveals important differences in situations when the ideal mode of a traveler is not the mode they actually use - or in other words, the affective dissonance between actual and ideal modes of travel. More concretely, the results indicate that users of public transportation experience the most dissonance, followed by automobile users, and finally active travelers, who tend to experience the least affective dissonance. For those travelers who experience dissonance, active travel is the mode most commonly associated with freedom, enjoyment, and happiness, whereas public transportation is most commonly associated with poverty, and the automobile is most commonly associated with luxury and status. We also find that there are significant variations in dissonance by age, education, income, and typical commute time. The reactions of travelers towards various transport modes are critical factors that policy-makers need to consider to promote and increase the use of public transport and active modes [@Bornioli2019affective; @Devos2017travel; @Devos2019satisfying; @Garling2019role; @Redman2013quality].

@@

As seen in the table, there are five individual attributes that we analyze. Three of these are socio-economic and demographic, namely age, level of education, and income. The other two are transportation related, namely primary mode of travel and commute time. The frequency tables are tested in every case using $\chi^2$ tests of independence ($p$-values are reported in the table; lower $p$-values mean that the null hypothesis of independence can be rejected with greater confidence). It is interesting to note that the only category for which all affective values are significant at better than 5% level of confidence is mode of travel.

@@

Turning now to typical commute time, four affective values show significant differences at better than 10% confidence: Freedom, Happiness, Luxury, and Status. Perhaps not surprisingly, dissonance is more frequent among people whose typical commutes are longer. This is in line with previous findings: both St-Louis et al. [-@Stlouis2014happy] and Smith [-@Smith2017commute] report that commute satisfaction tends to decline with longer commutes, whereas Handy and Thigpen [-@Handy2019commute] found that commute distance was a negative covariate of commute satisfaction.

@@

Attribute_by_ideal <- Modes_by_Attribute %>% 
  drop_na() %>%
  filter(ATTRIBUTE %in% c("FREEDOM", "ENJOYMENT", "HAPPINESS", "POVERTY", "LUXURY", "STATUS"),
         USED != "Other", 
         IDEAL != "Other") %>%
  group_by(IDEAL, ATTRIBUTE, USED) %>%
  summarize(n= sum(WEIGHT)) %>%
  ungroup() %>%
  # rename(MODE = IDEAL) %>%
  mutate(CASE = "IDEAL",
         USED = factor(USED, 
                       levels = c("Car", "Active", "Public"),
                       labels = c("USED: CAR", "USED: ACTIVE", "USED: PUBLIC")))

Attribute_by_used <- Attribute_by_ideal %>%
  group_by(USED, ATTRIBUTE) %>%
  summarize(t = sum(n))

@@

Attribute_by_ideal_by_mode <- ggplot(data = Attribute_by_ideal,
       aes(x = IDEAL, y = p, fill = IDEAL)) + 
  geom_col() +
  facet_grid(USED ~ ATTRIBUTE) +
  #xlab("Modes associated with affective values") +
  scale_y_continuous(name = "Proportion", 
                     breaks = c(0, 0.5, 1), 
                     limits = c(0, 1)) +
  theme_tufte() +
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "bottom",
        panel.border = element_rect(color = "black", fill = NA),
        axis.title.x = element_blank(),
        legend.title = element_blank())

@@

```{r figure-bar-plots-by-attribute-and-age, echo=FALSE, cache = TRUE, fig.cap="\\label{fig:bar-plots-by-age}Plots for affective values by age; in the y-axis is the proportion in the interval [0, 1], and the x-axis is the mode selected for each value (p-values are for Cochran-Mantel-Haenszel Chi-Squared Test)"}
Attribute_by_ideal_by_age <- Modes_by_Attribute %>% 
  drop_na() %>%
  filter(ATTRIBUTE %in% c("FREEDOM", "ENJOYMENT", "HAPPINESS", "POVERTY", "LUXURY", "STATUS"),
         USED != "Other", 
         IDEAL != "Other") %>%
  group_by(IDEAL, ATTRIBUTE, AGE, USED) %>%
  summarize(n= sum(WEIGHT)) %>%
  ungroup() %>%
  # rename(MODE = IDEAL) %>%
  mutate(CASE = "IDEAL",
         USED = factor(USED, 
                       levels = c("Car", "Active", "Public", "Other"),
                       labels = c("USED: CAR", "USED: ACTIVE", "USED: PUBLIC", "USED: OTHER")))

Attribute_by_ideal <- Attribute_by_ideal_by_age %>%
  group_by(USED, ATTRIBUTE, AGE) %>%
  summarize(t = sum(n))

@@

                                    axis.ticks.y=element_blank()),

@@

                                      axis.ticks.y=element_blank()),

@@

                                   axis.ticks.y=element_blank()),
             textGrob("Car",

@@

             textGrob("Active",
             textGrob("Public",

@@

                      y = 0.7),

@@

Attribute_by_ideal_by_education <- Modes_by_Attribute %>% 
  drop_na() %>%
  filter(ATTRIBUTE %in% c("FREEDOM", "ENJOYMENT", "HAPPINESS", "POVERTY", "LUXURY", "STATUS"),
         USED != "Other", 
         IDEAL != "Other") %>%
  group_by(IDEAL, ATTRIBUTE, EDUCATION, USED) %>%
  summarize(n= sum(WEIGHT)) %>%
  ungroup() %>%
  # rename(MODE = IDEAL) %>%
  mutate(CASE = "IDEAL",
         USED = factor(USED, 
                       levels = c("Car", "Active", "Public", "Other"),
                       labels = c("USED: CAR", "USED: ACTIVE", "USED: PUBLIC", "USED: OTHER")))

Attribute_by_ideal <- Attribute_by_ideal_by_education %>%
  group_by(USED, ATTRIBUTE, EDUCATION) %>%
  summarize(t = sum(n))

@@

                                   axis.ticks.y=element_blank()),
             textGrob("Car",

@@

             textGrob("Active",
             textGrob("Public",

@@

                      y = 0.7),

@@

Attribute_by_ideal_by_income <- Modes_by_Attribute %>% 
  drop_na() %>%
  filter(ATTRIBUTE %in% c("FREEDOM", "ENJOYMENT", "HAPPINESS", "POVERTY", "LUXURY", "STATUS"),
         USED != "Other", 
         IDEAL != "Other") %>%
  group_by(IDEAL, ATTRIBUTE, INCOME, USED) %>%
  summarize(n= sum(WEIGHT)) %>%
  ungroup() %>%
  # rename(MODE = IDEAL) %>%
  mutate(CASE = "IDEAL",
         USED = factor(USED, 
                       levels = c("Car", "Active", "Public", "Other"),
                       labels = c("USED: CAR", "USED: ACTIVE", "USED: PUBLIC", "USED: OTHER")))

Attribute_by_ideal <- Attribute_by_ideal_by_income %>%
  group_by(USED, ATTRIBUTE, INCOME) %>%
  summarize(t = sum(n))

@@

                                    axis.ticks.y=element_blank()),

@@

                                   axis.ticks.y=element_blank()),
             textGrob("Car",

@@

Among car users, Freedom is associated with the car in the case of every commute but the longest (>60 min), in which case active travel is more frequently identified as the mode that evokes feelings of Freedom. The associations of active travel with Enjoyment  and Happiness are consistently high among car users, although car becomes more frequently associated with this value for longer commutes. As we have seen before, active travelers seldom experience dissonance with respect to Freedom, Enjoyment, and Happiness, and if anything the strength of this association increases with longer commutes. Public transportation users again are split in their preference towards active travel and car as the modes more likely to evoke feelings of Freedom, Enjoyment, and Happiness, and there are only relatively minor differences in these preferences by length of commute. The last affective value that displays significant differences in dissonance by length of typical commute is Status. The car tends to dominate as the mode more frequently associated with this value among car and public transportation users. Active travelers, on the other hand, associate status more strongly with short or very long commutes: only when commutes are shorter than 20 minutes or longer than 60 minutes do active travelers associate feelings of Status with the car.

@@

Attribute_by_ideal_by_time <- Modes_by_Attribute %>% 
  drop_na() %>%
  filter(ATTRIBUTE %in% c("FREEDOM", "ENJOYMENT", "HAPPINESS", "POVERTY", "LUXURY", "STATUS"),
         USED != "Other", 
         IDEAL != "Other") %>%
  group_by(IDEAL, ATTRIBUTE, TRAVEL_TIME, USED) %>%
  summarize(n= sum(WEIGHT)) %>%
  ungroup() %>%
  # rename(MODE = IDEAL) %>%
  mutate(CASE = "IDEAL",
         USED = factor(USED, 
                       levels = c("Car", "Active", "Public", "Other"),
                       labels = c("USED: CAR", "USED: ACTIVE", "USED: PUBLIC", "USED: OTHER")))

Attribute_by_ideal <- Attribute_by_ideal_by_time %>%
  group_by(USED, ATTRIBUTE, TRAVEL_TIME) %>%
  summarize(t = sum(n))

@@

grid.arrange(textGrob(paste('FREEDOM\np-value', 
                            ifelse(round(Modes_by_Freedom_by_Time$p.value, 2) >= 0.001,
                                   paste("=", round(Modes_by_Freedom_by_Time$p.value, 3)),
                                   " < 0.001")),
                      gp = gpar(fontsize = 9)),
             textGrob(paste('ENJOYMENT\np-value', 
                            ifelse(round(Modes_by_Enjoyment_by_Time$p.value, 3) >= 0.001,
                                   paste("=", round(Modes_by_Enjoyment_by_Time$p.value, 3)),
                                   " < 0.001")),
                      gp = gpar(fontsize = 9)),
             textGrob(paste('HAPPINESS\np-value', 
                            ifelse(round(Modes_by_Happiness_by_Time$p.value, 3) >= 0.001,
                                   paste("=", round(Modes_by_Happiness_by_Time$p.value, 3)),
                                   " < 0.001")),
                      gp = gpar(fontsize = 9)),
             textGrob(paste('STATUS\np-value', ifelse(round(Modes_by_Status_by_Time$p.value, 3) >= 0.001,
                                                      paste("=", round(Modes_by_Status_by_Time$p.value, 3)),
                                                      " < 0.001")),
                      gp = gpar(fontsize = 9)),
             Freedom_by_time + theme(legend.position="none", 
                                    strip.text.x = element_text(angle = 90),
                                    strip.text.y = element_blank(),
                                    axis.title.y=element_blank(),
                                    axis.text.y=element_blank(),
                                    axis.ticks.y=element_blank()),
             Enjoyment_by_time + theme(legend.position="none", 
                                      strip.text.x = element_text(angle = 90),
                                      strip.text.y = element_blank(),
                                      axis.title.y=element_blank(),
                                      axis.text.y=element_blank(),
                                      axis.ticks.y=element_blank()),
             Happiness_by_time + theme(legend.position="none", 
                                      strip.text.x = element_text(angle = 90),
                                      strip.text.y = element_blank(),
                                      axis.title.y=element_blank(),
                                      axis.text.y=element_blank(),
                                      axis.ticks.y=element_blank()),
             Status_by_time + theme(legend.position="none",
                                   strip.text.x = element_text(angle = 90),
                                   strip.text.y = element_blank(),
                                   axis.title.y=element_blank(),
                                   axis.text.y=element_blank(),
                                   axis.ticks.y=element_blank()),
             textGrob("Car",
                      rot = 270,
                      gp = gpar(fontsize = 9),
                      y = 0.2),
             textGrob("Active",
                      rot = 270,
                      gp = gpar(fontsize = 9)),
             textGrob("Public",
                      rot = 270,
                      gp = gpar(fontsize = 9),
                      y = 0.7),
             mylegend, 
             layout_matrix = lay,
             #nrow=3,
             #widths = c(2.4, 1.9, 1.9, 1.9, 1.9), 
             heights = c(1, 3, 3, 3, 1),
             widths = c(3, 3, 3, 3, 0.5),
             left = "Proportion",
             right = "Primary Mode of Travel",
             top = "Affective Values by Commute Time",
             bottom = "Modes Associated with Affective Values")
```

>>>>

---
title: Do Drivers Dream of Walking? An Investigation of Travel Mode Dissonance from the Perspective of Affective Values
author:
  - name: Author 1
    email: a1@university.com
    affiliation: Some University
    footnote: Corresponding Author
  - name: Author 2
    email: a2@school.com
    affiliation: Some School
address:
  - code: Some University
    address: Department, Street, City, State, Zip
  - code: Some School
    address: Department, Street, City, State, Zip
abstract: |
   *Introduction*  
   Subjective wellbeing is a topic that has attracted considerable attention in the transportation literature in recent years. As a result, there is a burgeoning literature that investigates the impacts of travel on subjective wellbeing, and how wellbeing, in turn, can influence behavior. An important aspect of subjective wellbeing are the affective reactions of people to their experiences.   
   *Objective*  
   The objective of this paper is to analyze the affective reactions of travelers with respect to various modes of transportation. In particular, we are interested in the potential for dissonance between primary mode of travel and the mode(s) of travel identified as evoking various affective reactions.  
   *Materials and Methods*  
   The study is based on data collected from a sample of travelers in the city of Santiago, in Chile. Participants in the study were asked about their usual mode of travel, and then were asked to name their ideal mode(s) of transportation from the perspective of various affective reactions. The reactions we investigate are associated with the values of Freedom, Enjoyment, Happiness, Poverty, Luxury, and Status. Analysis is based on tests of independence and visualization techniques.  
   *Results*  
   The results indicate that users of public transportation experience the most dissonance in terms of affective reactions, and active travelers the least. For those travelers who experience dissonance, active travel is the mode most commonly associated with Freedom, Enjoyment, and Happiness, while public transportation is most commonly associated with Poverty. The automobile, in contrast, is the mode most commonly associated with Luxury and Status.

@@

# The master dataframe is processed for analysis. We plan to conduct two distinct types of analysis:
#
# 1. Who experiences dissonace?
#
# Analysis of dissonance using the characteristics of respondents (i.e., AGE, EDUCATION, INCOME, (Mode) USED, and TRAVEL_TIME) by ATTRIBUTE (i.e., FREEDOM et al.). In this case, dissonance is defined as follows:
#
# >> If the mode used is identical to the mode that the respondent associates with ATTRIBUTE (i.e. FREEDOM et al.), we consider that there is no dissonance; otherwise, we consider that there is dissonance. For example, if the respondent's main mode of transportation is Car, and the respondent associates Car with feelings of FREEDOM, then there is no dissonance. On the other hand, if the respondent's main mode of transportation is Car, but associates Walking with feelings of FREEDOM, there is dissonance.
# 
# Given this, we will calculate how frequently there is dissonance at different levels of AGE, EDUCATION, INCOME, MODE, TRAVEL_TIME. Tests of independence can be implemented.
#
# 2. For those respondents who experience dissonance, what mode(s) are better aligned with various affective values
#
# Analysis of dissonance using the mode USED and the mode respondents associated with ATTRIBUTE, then stratified by characteristics of respondents (i.e., AGE et al.) in this case, we see how frequently each mode was associated with ATTRIBUTE, given the main mode of transportation of the respondents.
#
# Here the dataframes are prepared to support these two types of analysis

@@

  ###################################################################################
#                                                                                 #
# Function created by Dr Paul Williamson, Dept. of Geography and Planning,        #
# School of Environmental Sciences, University of Liverpool, UK.                  #
#                                                                                 #
# Adapted from the function ctab() in the catspec package.                         #
#                                                                                 #
# Version: 12th July 2013                                                         #
#                                                                                 #
# Output best viewed using the companion function print.crosstab()                #
#                                                                                 #
###################################################################################

@@

#If no type supplied, default to 'frequency + total' for univariate tables and to
#'frequency' for multi-dimensional tables

@@

#If dec.places not set by user, set to 2 unless only one table of type frequency requested,
#in which case set to 0.  [Leaves user with possibility of having frequency tables with > 0 dp]
if (is.null(dec.places)) {
  if ((length(type)==1) & (type[1]=="frequency")) {
    dec.places <- 0
  } else {
    dec.places <-2
  }
}

@@

#If subtotals not required, and total vars > 2, create dataframe version of table, with relevant
#subtotal rows / cols dropped [Subtotals only present in tables with > 2 cross-classified vars]
t1 <- NULL
if ( (subtotals==FALSE) & (n.vars>2) )  {

@@

  #Remove backslashes from category names AND colnames
  t1 <- apply(t1[,],2, function(x) gsub("\"","",x))
  #Remove preceding and trailing spaces from category names to enable accurate capture of 'sum' rows/cols
  #[Use of grep might extract category labels with 'sum' as part of a longer one or two word string...]
  t1 <- apply(t1,2,function(x) gsub("[[:space:]]*$","",gsub("^[[:space:]]*","",x)))

@@

A definition of subjective wellbeing is as “[g]ood mental states, including all of the various evaluations, positive and negative, that people make of their lives, and the affective reactions of people to their experiences” [@oecd2013oecd, p. 10]. As memorably put by Steg [-@Steg2005car]: is the use of a car a must, or is it a lust? A close alignment, or consonance, between affective reactions and the mode of transportation used to travel can result in greater subjective wellbeing and increase the probability of choosing that mode (i.e., by satisfying the _lust_); on the other hand, dissonance, namely the lack of concordance between the mode used and the affective reaction to that mode, would be detrimental to subjective wellbeing, especially to those for whom the use of the mode is a _must_ [@Devos2019satisfaction; @Mokhtarian2018travel]. With these considerations in mind, the present research aims to investigate _who_ experiences affective dissonance with respect to their primary mode of travel. We focus our investigation on reactions from the perspective of feelings of Freedom, Enjoyment, Happiness, Poverty, Luxury, and Status - concepts termed _affective values_ hereafter. Furthermore, it is possible, even if use of a particular mode is a _must_, that travelers may still _lust_ for something else - in other words, the grass may look greener from the window of a car. For this reason, we also aim to investigate which mode or modes are most commonly identified as ideal from the perspective of the aforementioned affective values.

Our research is based on data collected from a sample of travelers in the city of Santiago in Chile. The analysis sheds light on the affective reactions of people towards their primary mode of travel, and their perceptions towards 'ideal modes'. This reveals important differences in situations when the ideal mode of a traveler is not the mode they actually use - or in other words, the affective dissonance between actual and ideal modes of travel. More concretely, the results indicate that users of public transportation experience the most dissonance, followed by automobile users, and finally active travelers, who tend to experience the least affective dissonance. For those travelers who experience dissonance, active travel is the mode most commonly associated with Freedom, Enjoyment, and Happiness, whereas public transportation is most commonly associated with Poverty, and the automobile is most commonly associated with Luxury and Status. We also find that there are significant variations in dissonance by age, education, income, and typical commute time. The reactions of travelers towards various transport modes are critical factors that policy-makers need to consider to promote and increase the use of public transport and active modes [@Bornioli2019affective; @Devos2017travel; @Devos2019satisfying; @Garling2019role; @Redman2013quality].

@@

As seen in the table, there are five individual attributes that we analyze. Three of these are socio-economic and demographic, namely age, level of education, and income. The other two are transportation related, namely primary mode of travel and commute time. The frequency tables are tested in every case using $\chi^2$ tests of independence ($p$-values are reported in the table; lower $p$-values mean that the null hypothesis of independence can be rejected with greater confidence). It is interesting to note that the only category for which all affective values are significant ($p\leq0.05$) is mode of travel.

@@

Turning now to typical commute time, four affective values show significant differences with $p\leq0.10$: Freedom, Happiness, Luxury, and Status. Perhaps not surprisingly, dissonance is more frequent among people whose typical commutes are longer. This is in line with previous findings: both St-Louis et al. [-@Stlouis2014happy] and Smith [-@Smith2017commute] report that commute satisfaction tends to decline with longer commutes, and Handy and Thigpen [-@Handy2019commute] found that commute distance was a negative covariate of commute satisfaction.

@@

Attribute_by_ideal <- Modes_by_Attribute %>% 
  drop_na() %>%
  filter(ATTRIBUTE %in% c("FREEDOM", "ENJOYMENT", "HAPPINESS", "POVERTY", "LUXURY", "STATUS"),
         USED != "Other", 
         IDEAL != "Other") %>%
  group_by(IDEAL, ATTRIBUTE, USED) %>%
  summarize(n= sum(WEIGHT),
            .groups = "drop") %>%
  ungroup() %>%
  # rename(MODE = IDEAL) %>%
  mutate(ATTRIBUTE = factor(ATTRIBUTE,
                            level = c("FREEDOM", "ENJOYMENT", "HAPPINESS", "POVERTY", "LUXURY", "STATUS")),
         CASE = "IDEAL",
         USED = factor(USED, 
                       levels = c("Car", "Active", "Public"),
                       labels = c("Used: Car", "Used: Active", "Used: Public")))

Attribute_by_used <- Attribute_by_ideal %>%
  group_by(USED, ATTRIBUTE) %>%
  summarize(t = sum(n),
            .groups = "drop")

@@

Attribute_by_ideal_by_mode <- ggplot(data = Attribute_by_ideal,
       aes(x = IDEAL, y = p, fill = IDEAL)) + 
  geom_col() +
  facet_grid(USED ~ ATTRIBUTE) +
  #xlab("Modes associated with affective values") +
  scale_y_continuous(name = "Proportion", 
                     breaks = c(0, 0.5, 1), 
                     limits = c(0, 1)) +
  theme_tufte() +
  theme(axis.text.x = element_blank(),
        legend.position = "bottom",
        panel.border = element_rect(color = "black", fill = NA),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        legend.title = element_blank())

@@

```{r figure-bar-plots-by-attribute-and-age, echo=FALSE, cache = TRUE, fig.cap="\\label{fig:bar-plots-by-age}Plots for affective values by age; in the y-axis is the proportion in the interval [0, 1], and the x-axis is the mode selected for each value (p-values are for Cochran-Mantel-Haenszel Chi-Squared Test)"}
Attribute_by_ideal_by_age <- Modes_by_Attribute %>% 
  drop_na() %>%
  filter(ATTRIBUTE %in% c("FREEDOM", "ENJOYMENT", "HAPPINESS", "POVERTY", "LUXURY", "STATUS"),
         USED != "Other", 
         IDEAL != "Other") %>%
  group_by(IDEAL, ATTRIBUTE, AGE, USED) %>%
  summarize(n= sum(WEIGHT),
            .groups = "drop") %>%
  ungroup() %>%
  # rename(MODE = IDEAL) %>%
  mutate(CASE = "IDEAL",
         USED = factor(USED, 
                       levels = c("Car", "Active", "Public", "Other"),
                       labels = c("USED: CAR", "USED: ACTIVE", "USED: PUBLIC", "USED: OTHER")))

Attribute_by_ideal <- Attribute_by_ideal_by_age %>%
  group_by(USED, ATTRIBUTE, AGE) %>%
  summarize(t = sum(n),
            .groups = "drop")

@@

                                    axis.title.y = element_blank(),
                                    axis.text.y = element_blank(),
                                    axis.ticks.y = element_blank(),
                                    axis.text.x = element_blank(),
                                    axis.ticks.x = element_blank()),

@@

                                      axis.title.y = element_blank(),
                                      axis.text.y = element_blank(),
                                      axis.ticks.y = element_blank(),
                                    axis.text.x = element_blank(),
                                    axis.ticks.x =element_blank()),

@@

                                   axis.title.y = element_blank(),
                                   axis.text.y = element_blank(),
                                   axis.ticks.y = element_blank(),
                                    axis.text.x = element_blank(),
                                    axis.ticks.x = element_blank()),
             textGrob("Used: Car",

@@

                      y = 0.1),
             textGrob("Used: Active",

@@

                      y = 0.3),
             textGrob("Used: Public",
                      y = 0.4),

@@

Attribute_by_ideal_by_education <- Modes_by_Attribute %>% 
  drop_na() %>%
  filter(ATTRIBUTE %in% c("FREEDOM", "ENJOYMENT", "HAPPINESS", "POVERTY", "LUXURY", "STATUS"),
         USED != "Other", 
         IDEAL != "Other") %>%
  group_by(IDEAL, ATTRIBUTE, EDUCATION, USED) %>%
  summarize(n= sum(WEIGHT),
            .groups = "drop") %>%
  ungroup() %>%
  # rename(MODE = IDEAL) %>%
  mutate(CASE = "IDEAL",
         USED = factor(USED, 
                       levels = c("Car", "Active", "Public", "Other"),
                       labels = c("USED: CAR", "USED: ACTIVE", "USED: PUBLIC", "USED: OTHER")))

Attribute_by_ideal <- Attribute_by_ideal_by_education %>%
  group_by(USED, ATTRIBUTE, EDUCATION) %>%
  summarize(t = sum(n),
            .groups = "drop")

@@

                                   axis.title.y = element_blank(),
                                   axis.text.y = element_blank(),
                                   axis.ticks.y = element_blank(),
                                   axis.text.x = element_blank(),
                                   axis.ticks.x = element_blank()),
             textGrob("Used: Car",
                      y = 0.0),
             textGrob("Used: Active",

@@

             textGrob("Used: Public",

@@

                      y = 0.4),

@@

Attribute_by_ideal_by_income <- Modes_by_Attribute %>% 
  drop_na() %>%
  filter(ATTRIBUTE %in% c("FREEDOM", "ENJOYMENT", "HAPPINESS", "POVERTY", "LUXURY", "STATUS"),
         USED != "Other", 
         IDEAL != "Other") %>%
  group_by(IDEAL, ATTRIBUTE, INCOME, USED) %>%
  summarize(n= sum(WEIGHT),
            .groups = "drop") %>%
  ungroup() %>%
  # rename(MODE = IDEAL) %>%
  mutate(CASE = "IDEAL",
         USED = factor(USED, 
                       levels = c("Car", "Active", "Public", "Other"),
                       labels = c("USED: CAR", "USED: ACTIVE", "USED: PUBLIC", "USED: OTHER")))

Attribute_by_ideal <- Attribute_by_ideal_by_income %>%
  group_by(USED, ATTRIBUTE, INCOME) %>%
  summarize(t = sum(n),
            .groups = "drop")

@@

                                    axis.ticks.y=element_blank(),
                                   axis.text.x = element_blank(),
                                   axis.ticks.x = element_blank()),

@@

                                   axis.ticks.y=element_blank(),
                                   axis.text.x = element_blank(),
                                   axis.ticks.x = element_blank()),
             textGrob("Used: Car",

@@

Among car users, Freedom is associated with the car in all but the longest commutes (>60 min), in which case active travel is more frequently identified as the mode that evokes feelings of Freedom. The associations of active travel with Enjoyment  and Happiness are consistently high among car users, although car becomes more frequently associated with this value for longer commutes. As we have seen before, active travelers seldom experience dissonance with respect to Freedom, Enjoyment, and Happiness, and if anything the strength of this association increases with longer commutes. Public transportation users again are split in their preference towards active travel and car as the modes more likely to evoke feelings of Freedom, Enjoyment, and Happiness, and there are only relatively minor differences in these preferences by length of commute. The last affective value that displays significant differences in dissonance by length of typical commute is Status. The car tends to dominate as the mode more frequently associated with this value among car and public transportation users. Active travelers, on the other hand, associate status more strongly with short or very long commutes: only when commutes are shorter than 20 minutes or longer than 60 minutes do active travelers associate feelings of Status with the car.

@@

Attribute_by_ideal_by_time <- Modes_by_Attribute %>% 
  drop_na() %>%
  filter(ATTRIBUTE %in% c("FREEDOM", "ENJOYMENT", "HAPPINESS", "POVERTY", "LUXURY", "STATUS"),
         USED != "Other", 
         IDEAL != "Other") %>%
  group_by(IDEAL, ATTRIBUTE, TRAVEL_TIME, USED) %>%
  summarize(n= sum(WEIGHT),
            .groups = "drop") %>%
  ungroup() %>%
  # rename(MODE = IDEAL) %>%
  mutate(CASE = "IDEAL",
         USED = factor(USED, 
                       levels = c("Car", "Active", "Public", "Other"),
                       labels = c("USED: CAR", "USED: ACTIVE", "USED: PUBLIC", "USED: OTHER")))

Attribute_by_ideal <- Attribute_by_ideal_by_time %>%
  group_by(USED, ATTRIBUTE, TRAVEL_TIME) %>%
  summarize(t = sum(n),
            .groups = "drop")

@@

                            ifelse(round(Modes_by_Freedom_by_Time$p.value, 2) >= 0.001,
                                   paste("=", round(Modes_by_Freedom_by_Time$p.value, 3)),
                            ifelse(round(Modes_by_Enjoyment_by_Time$p.value, 3) >= 0.001,
                                   paste("=", round(Modes_by_Enjoyment_by_Time$p.value, 3)),
                            ifelse(round(Modes_by_Happiness_by_Time$p.value, 3) >= 0.001,
                                   paste("=", round(Modes_by_Happiness_by_Time$p.value, 3)),
             textGrob(paste('STATUS\np-value', ifelse(round(Modes_by_Status_by_Time$p.value, 3) >= 0.001,
                                                      paste("=", round(Modes_by_Status_by_Time$p.value, 3)),
             Freedom_by_time + theme(legend.position="none", 
                                    axis.ticks.y=element_blank(),
                                   axis.text.x = element_blank(),
                                   axis.ticks.x = element_blank()),
             Enjoyment_by_time + theme(legend.position="none", 
                                      axis.ticks.y=element_blank(),
                                   axis.text.x = element_blank(),
                                   axis.ticks.x = element_blank()),
             Happiness_by_time + theme(legend.position="none", 
                                      axis.ticks.y=element_blank(),
                                   axis.text.x = element_blank(),
                                   axis.ticks.x = element_blank()),
             Status_by_time + theme(legend.position="none",
                                   axis.ticks.y=element_blank(),
                                   axis.text.x = element_blank(),
                                   axis.ticks.x = element_blank()),
             textGrob("Used: Car",
                      y = 0.0),
             textGrob("Used: Active",
             textGrob("Used: Public",
                      y = 0.4),
             widths = c(3, 3, 3, 3, 0.5),
             top = "Affective Values by Commute Time",
